null

### 2017-03-31 Manuscript/Manuscript_COD_Trends.Rmd
Update Manuscript_COD_Trends.Rmd
Hey Corinne-
I have not made a full pass through this yet, but tried to just give a couple of comments on the Intro and Results. Generally speaking, I think this the right way to go in terms of describing the results--report on the general pattern among all states but identify some key outliers of interest, particularly if they are regional in nature, which could be linked to other changes in risk factors, etc.

The over-arching objective of this study was to describe the general pattern of each cause of death's contribution to the black-white life expectancy gap over forty-five years, and identify divergent states and differences between the sexes. To do so, we estimated state-level trends in the contribution of six major causes of death to the black-white life expectancy gap for males and females. Common reasons for the decreasing life expectancy gap can thereby be identified, and anomalous states can be examined. Ideally, this study may facilitate the identification of one policies or social conditions in highly successful states, and suggest broad targets for interventions to further reduce the gap in less successful states.

@@

**Figures 1 and 2** display trends in the life expectancy gap for males and females, respectively. For both sexes, the life expectancy gap has decreased since 1969, although several states exhibited a period of increasing inequality that peaked in the early 1990s, especially among males.

**Figures 2 - 14** display trends in the contribution of each cause of death to the life expectancy gap for males and females, in years. The blue band illustrates the credible interval for the smoothed contribution estimated for each state, and illustrates the heterogeneity across states in their contributions trends. These curves are overlaid with the meta-estimated common trend, shown with a red line, and the plots are organized according to the closeness of the [fit] (*need another word*) of the red curve to the blue band. Thus, increasingly divergent states are shown further down the display. We supplement these main results with smoothed estimates of the age-standardized mortality rates for each cause of death (**eFigures 1-12**).

@@

For females, while cancer-related mortality continues to be higher among blacks in many states, the inverted-U pattern is similar between blacks and whites and most states do not exhibit the accentuated period of increase in mortality that was estimated for black males. Thus, the contribution of cancer to the life expectancy gap is generally flat over time (**Figure 6**). Washington DC deviates from this pattern, and exhibited increasing cancer contributions to the gap, while Rhode Island exhibited the reverse. Most notably, some states exhibited no inequality in cancer mortality (**eFigures 3 and 4**), including New York, and Connecticut (and for males, these states recently eliminated the black-white difference in cancer mortality). Furthermore, Massachusetts and (imprecisely) Rhode Island exhibit lower cancer rates among black women and men compared to whites in 2013.

>>>>

The over-arching objective of this study was to describe the pattern of each cause of death's contribution to the black-white life expectancy gap over forty-five years, and identify divergent states and gender differences. To do so, we estimated state-level trends in the contribution of six major causes of death to the black-white life expectancy gap for males and females. This facilitates the identification of common reasons for the decreasing life expectancy gap and the identification of anomalous states. Ideally, this study may facilitate the identification of one policies or social conditions in highly successful states, and suggest broad targets for interventions to further reduce the gap in less successful states.

@@

**Figures 1 and 2** display trends in the life expectancy gap for males and females, respectively. For both sexes, the life expectancy gap has decreased since 1969, although several states exhibited a period of increasing inequality that peaked in the early 1990s, especially among males. Don't know whether you want to note it here or not, but I find it is interesting to see that the gap is now lower in many US southern (or southeastern) states than in the Midwest, suggesting that progress has actually been somewhat greater in the South than elsewhere (particularly the upper Midwest and especially for women). I think this is interesting and somewhat counter to the narrative that health indicators are worse in the South. I also wonder how this might relate to the Murray/IHME work on US counties with "declining" life expectancya among white women.

**Figures 2 - 14** display trends in the contribution of each cause of death to the life expectancy gap for males and females, in years. [I wonder whether some readers might be confused by the changes in the ordering of the plots from the geographic layout in F1/2 to the rest of the plots]. The blue band illustrates the credible interval for the smoothed contribution estimated for each state, and illustrates the heterogeneity across states in their contributions trends. These curves are overlaid with the meta-estimated common trend, shown with a red line, and the plots are organized according to the closeness of the [fit] (*need another word*) of the red curve to the blue band. Thus, increasingly divergent states are shown further down the display. We supplement these main results with smoothed estimates of the age-standardized mortality rates for each cause of death (**eFigures 1-12**).

@@

For females, while cancer-related mortality continues to be higher among blacks in many states, the inverted-U pattern is similar between blacks and whites and most states do not exhibit the accentuated period of increase in mortality that was estimated for black males. Thus, the contribution of cancer to the life expectancy gap is generally flat over time (**Figure 6**). [You could potentially stop here to save space and just include the comment about there being little inequality in cancer mortality ]. Washington DC deviates from this pattern, and exhibited increasing cancer contributions to the gap, while Rhode Island exhibited the reverse. Most notably, some states exhibited no inequality in cancer mortality (**eFigures 3 and 4**), including New York, and Connecticut (and for males, these states recently eliminated the black-white difference in cancer mortality). Furthermore, Massachusetts and (imprecisely) Rhode Island exhibit lower cancer rates among black women and men compared to whites in 2013.
